An Embedding Framework for Consistent Polyhedral Surrogates

### 2022-05-11 poly-embed-journal/poly-embed-journal.tex
meeting notes from yesterday

\begin{definition}
  \label{def:calibrated}
  Let discrete loss $\ell:\R\to\reals^\Y_+$, proposed surrogate $L:\reals^d\to\reals^\Y_+$, and link function $\psi:\reals^d\to\R$ be given.
  We say $(L,\psi)$ is \emph{calibrated} with respect to $\ell$ if
for all $p \in \simplex$,
  \begin{equation}
    \label{eq:calibrated}
  \inf_{u \in \reals^d : \psi(u) \not\in \prop{\ell}(p)} \inprod{p}{L(u)} > \inf_{u \in \reals^d} \inprod{p}{L(u)}~.
  \end{equation}
  If $(L, \psi)$ is calibrated with respect to $\ell$, we call $\psi$ a \emph{calibrated link.}
\end{definition}
It is well-known in finite-outcome settings that calibration is equivalent to \emph{consistency}, in the following sense (cf.~\citep{bartlett2006convexity,zhang2004statistical,agarwal2015consistent}).
Suppose we have the feature space $\X$ and label space $\Y$.
%For any data distribution $D \in \Delta(\X \times \Y)$, let $L^*$ be the best possible expected (over $D$) $L$-loss achieved by any hypothesis $H:\X\to\reals^d$, and $\ell^*$ the best expected $\ell$-loss for any hypothesis $h:\X\to\R$, respectively.
We say a surrogate and link pair $(L,\psi)$ is consistent with respect to $\ell$ if, for all data distributions $D \in \Delta(\X \times \Y)$, and all sequences of surrogate hypotheses $H_1,H_2,\ldots$ whose $L$-loss limits to the optimal surrogate loss $L^*$ (in expectation over $D$), the $\ell$-loss of the sequence $\psi\circ H_1,\psi \circ H_2, \ldots$ limits to the optimal target loss $\ell^*$.
\bo{edited above slightly. I believe a technical definition of consistency would have a ``for all $\D$'', but we can probably skip that.}\jessie{Tried to make sure it had this technical correctness.}
\raf{Tried to clarify/correct the quantifiers.  OLD VERSION:
Let $L^*$ be the best possible expected $L$-loss achieved by any hypothesis $H:\X\to\reals^d$, and $\ell^*$ the best expected $\ell$-loss for any hypothesis $h:\X\to\R$, respectively.
If $(L,\psi)$ is consistent, then for any sequence of surrogate hypotheses $H_1,H_2,\ldots$ whose $L$-loss limits to $L^*$, the $\ell$-loss of $\psi\circ H_1,\psi \circ H_2, \ldots$ limits to $\ell^*$ for all data distributions $D \in \Delta(\X \times \Y)$.}
%
As Definition~\ref{def:calibrated} does not involve the feature space $\X$, we will drop it for the remainder of the paper.
\jessiet{Commented out the following sentence as it was repetitive from earlier discussion. - Mar 24 22}

Like the use of a surrogate and link pair in the calibration definition, one can also extend the earlier definition of property elicitation to \emph{indirect (property) elicitation}, in which one applies a link to an elicited property to obtain a related property of interest.
\begin{definition}\label{def:indirect-elic}
	A loss $L : \reals^d \to \reals^\Y_+$  and surrogate $\psi : \reals^d \to \R$ pair $(L, \psi)$ indirectly elicits a property $\gamma : \simplex \toto \R$ if $L$ elicits a property $\Gamma : \simplex \toto \reals^d$ such that, for all $u \in \reals^d$, we have $\Gamma_u \subseteq \gamma_{\psi(u)}$.
	Moreover, we say $L$ indirectly elicits $\gamma$ if such a $\psi$ exists.
\end{definition}

At times, the elicitation literature studies indirect elicitation through the notion of one property ``refining'' another~\citep{frongillo2015elicitation}, where a property $\Gamma: \simplex \toto \R$ refines $\Gamma': \simplex \toto \R'$ if, for all $r \in \R$, there exists an $r' \in \R'$ such that $\Gamma_r \subseteq \Gamma'_{r'}$.
From the definition of refinement, it is almost immediate to observe that a loss $L$ indirectly elicits a property $\gamma$ if $\prop L$ refines $\gamma$.

@@

\subsection{Applying the embedding framework}
When using our framework to study the (in)consistency of an existing surrogate $L:\reals^d \to \reals^\Y_+$, often the first step is determining the loss it embeds.
To this end, we suggest the following general approach.
First, for each $y\in\Y$, divide $\reals^d$ into a finite number of polyhedral regions on which $L(\cdot)_y$ is an affine function.
Second, identify the vertices of these polyhedral regions.%
\footnote{In some cases, these regions do not have vertices, such as the top-$k$ surrogates which are invariant in the all-ones direction; here one can restrict to a subspace, or otherwise select among equivalent reports.}
Third, conclude that the union of these vertices, $\Sc\subset\reals^d$, is a finite representative set for $L$.
% ~\citep[Statement 3.1.7]{grunbaum2013convex}.
Now $L$ embeds $L|_\Sc$ from Proposition~\ref{prop:representative-embeds-restriction}.
From here one can further remove redundant reports until arriving at a tight embedding if desired.
Once the embedded discrete loss is known, the behavior of the surrogate becomes more clear: what problem it is solving, and for which restrictions on label distributions is it consistent for the original problem.

%% giving intuition for new figure
\raf{Follow-up paragraph on prev sentence using abstain as an example} \jessie{Attempted}
For example, \citet{ramaswamy2018consistent} propose the BEP surrogate, which embeds the discrete abstain loss (\S~\ref{sec:abstain}).
With the abstain loss in hand, it is easy to verify that the abstain-optimal report for any $p$ such that $\max_y p_y < 1/2$ is to ``abstain'' from predicting.
When juxtaposed with multiclass classification in Figure~\ref{fig:abstain-intuition-inconsistent}, one can observe that the BEP surrogate does not indirectly elicit the mode, as the level set $\prop{L^{BEP}}_{\vec 0} = \prop{\ell^{\mathrm{abs}}}_\bot \not \subseteq \mode_r$ for any $r$. \jessie{Notation}
As embedding is a special form of indirect elicitaion, we additionally know the BEP surrogate does not embed the mode.
The intuition described by Figure~\ref{fig:abstain-intuition-inconsistent} additionally gives a procedure to validate inconsistency with respect to a target task.

@@

Using our framework, one can show that $L^{1/2}$ embeds (2 times) $\ell^{1/2}$, with the embedding given by $\varphi$ above where we define $\varphi(\bot) = 0 \in \reals^d$.
(Following the general procedure outlined above, the regions where $L^{1/2}$ is affine all have vertices in the set $\{-1,1\}^d \cup \{0\}$, meaning it is representative, and $L^{1/2}$ restricted to that set is precisely $2\ell^{1/2} \circ \varphi^{-1}$.)
%In light of our framework, we can see that $L^{1/2}$ is an excellent example of an embedding, where $\varphi(y) = B(y)$ and $\varphi(\bot) = 0 \in \reals^d$.
Moreover, as we illustrate in Figure~\ref{fig:abstain-links}(L), the link $\psi^{\epsilon}$ proposed by \citeauthor{ramaswamy2018consistent} can be recovered from Construction~\ref{const:eps-thick-link} by choosing the norm $\|\cdot\|_\infty$ for any sufficiently small choice of $\epsilon > 0$.
% and choosing $\psi^{1/2}(u) = \bot$ for all $u$ such that $\bot \in \Psi(u)$. 
%(For all $u$ such that $\bot \not \in \Psi(u)$, there is only one choice to link, e.g., $|\Psi(u)| = 1$.)
Hence, our framework could have simplified the process of finding $\psi^{1/2}$, and the corresponding proof of consistency and surrogate regret bounds (\S~\ref{subsec:regret-bounds}\raft{check whether our bounds match}).
% can be used to verify the calibration of the prposed link.
%\jessiet{changed this sentence.  Thoughts?}\raft{Changed it back, since I think our narrative here is that our framework would have simplified their paper}
To illustrate this point further, we give an alternate link $\psi^1$ corresponding to $\|\cdot\|_1$ and $\epsilon=1$, shown in Figure~\ref{fig:abstain-links}(R),
\begin{equation}\label{eq:abstain-link-1}
  \psi^1(u) = \begin{cases}
	\bot & \|u\|_1 \leq 1\\
	\varphi^{-1}(\sgn(u)) &\text{otherwise}
  \end{cases}~.
\end{equation}
Construction~\ref{const:eps-thick-link} gives calibration of $(L^{1/2},\psi^1)$ with respect to $\ell^{1/2}$. 
%\jessiet{does it?  do we know $\epsilon$ is small enough for higher $d$?}\raft{removed ``immediately'' since you're right there are details to work out, but I'm pretty sure $\epsilon=1$ should hold for all $d$}
Aside from its simplicity, one possible advantage of $\psi^1$ is that it assigns $\bot$ to much less of the surrogate space $\reals^d$\raft{Bo, we used to say ``it appears to yield the same constant in generalization bounds as $\psi$''but I think our paper implies that the bound is worse, right? EDIT 3/16/2022: if we add surrogate regret bounds to the paper, I think it could be fun and worth working out which link actually gives the better bound, at least from our best analysis.}

@@

Lemma~\ref{lem:X} leads to another appealing equivalent condition to our embedding condition in Definition~\ref{def:loss-embed}: a surrogate embeds a discrete loss if and only if their Bayes risks match.
\jessiet{Would it be fair to write about how this could give a quick emprical test of embedding by comparing Bayes risks at a few sampled distributions?}

@@

\subsection{Trimming a loss}
\jessiet{Should this subsection be merged with the one below?}
\jessiet{Ordering suggestion: non-redundancy, trimming, (cor relating non-redundancy and trim), trimming related to properties.}

@@

@@

%% first step to show iff: indirect polyhedral elicitation passes through embedding.
A useful lemma is that for minimizable polyhedral losses, indirect elicitation must always pass through an embedding.
This result holds more generally whenever $L$ has a finite representative set, as in \S~\ref{sec:min-rep-sets}.
\begin{lemma}\label{lem:ie-iff-embeds-refinement}
  Let $L$ be a minimizable polyhedral loss.
  Then $L$ indirectly elicits a property $\gamma$ if and only if $L$ tightly embeds a discrete loss $\ell$ such that $\prop \ell$ refines $\gamma$.
\end{lemma}
\begin{proof}
  Let $L:\reals^d\to\reals^\Y_+$ be polyhedral, and $\Gamma = \prop L$.
  Then $L$ tightly embeds a discrete loss from Lemma~\ref{lem:X}(\ref{item:X-tight-embed}).
  Furthermore, Lemma~\ref{lem:X}(\ref{item:X-min-Theta},\ref{item:X-redundant},\ref{item:X-tight-embed}) implies that $\prop L$ refines $\prop \ell$ for any discrete loss $\ell$ that $L$ tightly embeds.

>>>>

\begin{definition}
  \label{def:calibrated}
  Let discrete loss $\ell:\R\to\reals^\Y_+$, proposed surrogate $L:\reals^d\to\reals^\Y_+$, and link function $\psi:\reals^d\to\R$ be given.
  We say $(L,\psi)$ is \emph{calibrated} with respect to $\ell$ if
for all $p \in \simplex$,
  \begin{equation}
    \label{eq:calibrated}
  \inf_{u \in \reals^d : \psi(u) \not\in \prop{\ell}(p)} \inprod{p}{L(u)} > \inf_{u \in \reals^d} \inprod{p}{L(u)}~.
  \end{equation}
  If $(L, \psi)$ is calibrated with respect to $\ell$, we call $\psi$ a \emph{calibrated link.}
\end{definition}
It is well-known in finite-outcome settings that calibration is equivalent to \emph{consistency}, in the following sense (cf.~\citep{bartlett2006convexity,zhang2004statistical,agarwal2015consistent}).
Suppose we have the feature space $\X$ and label space $\Y$.
%For any data distribution $D \in \Delta(\X \times \Y)$, let $L^*$ be the best possible expected (over $D$) $L$-loss achieved by any hypothesis $H:\X\to\reals^d$, and $\ell^*$ the best expected $\ell$-loss for any hypothesis $h:\X\to\R$, respectively.
We say a surrogate and link pair $(L,\psi)$ is consistent with respect to $\ell$ if, for all data distributions $D \in \Delta(\X \times \Y)$, and all sequences of surrogate hypotheses $H_1,H_2,\ldots$ whose $L$-loss limits to the optimal surrogate loss $L^*$ (in expectation over $D$), the $\ell$-loss of the sequence $\psi\circ H_1,\psi \circ H_2, \ldots$ limits to the optimal target loss $\ell^*$.
As Definition~\ref{def:calibrated} does not involve the feature space $\X$, we will drop it for the remainder of the paper.

Like the use of a surrogate and link pair in the calibration definition, one can also extend the earlier definition of property elicitation to \emph{indirect (property) elicitation}, in which one applies a link to an elicited property to obtain a related property of interest.
\begin{definition}\label{def:indirect-elic}
	Let minimizable loss $L : \reals^d \to \reals^\Y_+$ and link $\psi : \reals^d \to \R$ be given.
  The pair $(L, \psi)$ indirectly elicits a property $\gamma : \simplex \toto \R$ if for all $u \in \reals^d$, we have $\Gamma_u \subseteq \gamma_{\psi(u)}$, where $\Gamma = \prop{L}$.
	Moreover, we say $L$ indirectly elicits $\gamma$ if such a $\psi$ exists.
\end{definition}

At times, the elicitation literature studies indirect elicitation through the notion of one property ``refining'' another~\citep{frongillo2015elicitation}, where a property $\Gamma: \simplex \toto \R$ refines $\Gamma': \simplex \toto \R'$ if, for all $r \in \R$, there exists an $r' \in \R'$ such that $\Gamma_r \subseteq \Gamma'_{r'}$.
In particular, a loss $L$ indirectly elicits a property $\gamma$ if and only if $\prop L$ refines $\gamma$.

@@

\subsection{Applying the embedding framework}
When using our framework to study the (in)consistency of an existing surrogate $L:\reals^d \to \reals^\Y_+$, often the first step is determining the loss it embeds.
To this end, we suggest the following general approach.
First, for each $y\in\Y$, divide $\reals^d$ into a finite number of polyhedral regions on which $L(\cdot)_y$ is an affine function.
Second, identify the vertices of these polyhedral regions.%
\footnote{In some cases, these regions do not have vertices, such as the top-$k$ surrogates which are invariant in the all-ones direction; here one can restrict to a subspace, or otherwise select among equivalent reports.}
Third, conclude that the union of these vertices, $\Sc\subset\reals^d$, is a finite representative set for $L$.
% ~\citep[Statement 3.1.7]{grunbaum2013convex}.
Now $L$ embeds $L|_\Sc$ from Proposition~\ref{prop:representative-embeds-restriction}.
From here one can further remove redundant reports until arriving at a tight embedding if desired.

Once the embedded discrete loss is known, the behavior of the surrogate becomes more clear: what problem it is solving, and for which restrictions on label distributions is it consistent for the original problem $\ell$.
In particular, any level set of the embedded property $\hat\gamma$ which spans multiple level sets of the target property $\gamma=\prop{\ell}$ will lead to inconsistency for $\ell$ (Figure~\ref{fig:abstain-intuition-inconsistent}).
To obtain consistency with respect to a desired target, therefore, it suffices to restrict to the union of level sets of $\hat\gamma$ which are each fully contained in some level set of $\gamma$.

@@

Using our framework, one can show that $L^{1/2}$ embeds (2 times) $\ell^{1/2}$, with the embedding given by $\varphi$ above where we define $\varphi(\bot) = 0 \in \reals^d$.
(Following the general procedure outlined above, the regions where $L^{1/2}$ is affine all have vertices in the set $\{-1,1\}^d \cup \{0\}$, meaning it is representative, and $L^{1/2}$ restricted to that set is precisely $2\ell^{1/2} \circ \varphi^{-1}$.)
%In light of our framework, we can see that $L^{1/2}$ is an excellent example of an embedding, where $\varphi(y) = B(y)$ and $\varphi(\bot) = 0 \in \reals^d$.

\raf{MOVED}%% giving intuition for new figure
With the abstain loss in hand, it is easy to verify that the abstain-optimal report for any $p$ such that $\max_y p_y < 1/2$ is to ``abstain'' from predicting.
When juxtaposed with multiclass classification in Figure~\ref{fig:abstain-intuition-inconsistent}, one can observe that the BEP surrogate does not indirectly elicit the mode, as the level set $\prop{L^{BEP}}_{\vec 0} = \prop{\ell^{\mathrm{abs}}}_\bot \not \subseteq \mode_r$ for any $r$. \jessie{Notation}
As embedding is a special form of indirect elicitaion, we additionally know the BEP surrogate does not embed the mode.
The intuition described by Figure~\ref{fig:abstain-intuition-inconsistent} additionally gives a procedure to validate inconsistency with respect to a target task.

Moreover, as we illustrate in Figure~\ref{fig:abstain-links}(L), the link $\psi^{\epsilon}$ proposed by \citeauthor{ramaswamy2018consistent} can be recovered from Construction~\ref{const:eps-thick-link} by choosing the norm $\|\cdot\|_\infty$ for any sufficiently small choice of $\epsilon > 0$.
% and choosing $\psi^{1/2}(u) = \bot$ for all $u$ such that $\bot \in \Psi(u)$. 
%(For all $u$ such that $\bot \not \in \Psi(u)$, there is only one choice to link, e.g., $|\Psi(u)| = 1$.)
Hence, our framework could have simplified the process of finding $\psi^{1/2}$, and the corresponding proof of consistency and surrogate regret bounds (\S~\ref{subsec:regret-bounds}\raft{check whether our bounds match}).
% can be used to verify the calibration of the prposed link.
%\jessiet{changed this sentence.  Thoughts?}\raft{Changed it back, since I think our narrative here is that our framework would have simplified their paper}
To illustrate this point further, we give an alternate link $\psi^1$ corresponding to $\|\cdot\|_1$ and $\epsilon=1$, shown in Figure~\ref{fig:abstain-links}(R),
\begin{equation}\label{eq:abstain-link-1}
  \psi^1(u) = \begin{cases}
	\bot & \|u\|_1 \leq 1\\
	\varphi^{-1}(\sgn(u)) &\text{otherwise}
  \end{cases}~.
\end{equation}
Construction~\ref{const:eps-thick-link} gives calibration of $(L^{1/2},\psi^1)$ with respect to $\ell^{1/2}$. 
%\jessiet{does it?  do we know $\epsilon$ is small enough for higher $d$?}\raft{removed ``immediately'' since you're right there are details to work out, but I'm pretty sure $\epsilon=1$ should hold for all $d$}
Aside from its simplicity, one possible advantage of $\psi^1$ is that it assigns $\bot$ to much less of the surrogate space $\reals^d$\raft{Bo, we used to say ``it appears to yield the same constant in generalization bounds as $\psi$''but I think our paper implies that the bound is worse, right? EDIT 3/16/2022: if we add surrogate regret bounds to the paper, I think it could be fun and worth working out which link actually gives the better bound, at least from our best analysis.}

@@

Lemma~\ref{lem:X} leads to another appealing equivalent condition to our embedding condition in Definition~\ref{def:loss-embed}: a surrogate embeds a discrete loss if and only if their Bayes risks match.

@@

\subsection{Trimming a loss}

@@

\raf{Short parag on why we care about trim}

@@

%% first step to show iff: indirect polyhedral elicitation passes through embedding.
\raf{Jessie: sync up refines vs indirect elic; use the latter unless it's really awkward}
A useful lemma is that for minimizable polyhedral losses, indirect elicitation must always pass through an embedding.
This result holds more generally whenever $L$ has a finite representative set, as in \S~\ref{sec:min-rep-sets}.
\begin{lemma}\label{lem:ie-iff-embeds-refinement}
  Let $L$ be a minimizable polyhedral loss.
  Then $L$ indirectly elicits a property $\gamma$ if and only if $L$ tightly embeds a discrete loss $\ell$ that indirectly elicits $\gamma$.
\end{lemma}
\begin{proof}
  Let $L:\reals^d\to\reals^\Y_+$ be polyhedral, and $\Gamma = \prop L$.
  Then $L$ tightly embeds a discrete loss from Lemma~\ref{lem:X}(\ref{item:X-tight-embed}).
  Furthermore, Lemma~\ref{lem:X}(\ref{item:X-min-Theta},\ref{item:X-redundant},\ref{item:X-tight-embed}) implies that $\prop L$ refines $\prop \ell$ for any discrete loss $\ell$ that $L$ tightly embeds.
